A Traveling Salesman Solution For The Capitals of All African Nations

### 2009-11-01 paper/paper.tex
Stuff
stuff
paper

\usepackage{graphicx}
\usepackage{multicol}
\usepackage{mdwlist}

@@

\begin{figure}
\begin{center}
\setlength\fboxsep{1.00pt}
\setlength\fboxrule{1.00pt}
\fbox{\includegraphics[scale=0.80]{points.png}}
\caption{Capitals of African Nations}
\end{center}
\end{figure}

@@

\section{Overview}
The remainder of this article is organized as follows.
Section~ gives account of previous work.
Our new and exciting results are described in Section~.
Finally, Section gives the conclusions.

\section{Programs}
A much longer  example was written by Gil

@@

\begin{figure}
\begin{center}
\setlength\fboxsep{1.00pt}
\setlength\fboxrule{1.00pt}
\fbox{\includegraphics[scale=0.65]{path.png}}
\caption{Final Path Through Africa}
\end{center}
\end{figure}

In this section we describe the results.

\section{Runtime}
We worked hard, and achieved very little.

@@

\end{document}

>>>>

\usepackage{graphicx}
\usepackage{multicol}
\usepackage{mdwlist}
\usepackage{amssymb,amsmath}
\usepackage{cite}
\usepackage[bookmarks=false,colorlinks=true,linkcolor={blue},pdfstartview={XYZ null null 1.22}]{hyperref}
\usepackage{url}
\usepackage[T1]{fontenc}
%\usepackage{gfsartemisia-euler}

@@

\begin{figure}[h!]
  \begin{center}
    \setlength\fboxsep{1.00pt}
    \setlength\fboxrule{1.00pt}
    \fbox{\includegraphics[scale=0.60]{points.png}}
    \caption{Capitals of African Nations}
  \end{center}
\end{figure}

@@

\section{Overview}

  The traveling salesman problem (TSP) is a thoroughly researched problem in theoreticaly computer-science. 
TSP is classified as NP-Hard eg. as hard or harder to solve than a problem solvable in nondeterministic polynomial time.
This means the worst case performance for a TSP algorithm will likely increase exponetially with the numbers of cities to traverse.

The big-O complexity of a brute force TSP algorithms (check all vertices against all other vertices) is $O(n!)$.

\section{Programs}

\begin{enumerate}
\item The Python 2.6 \cite{python} programming language and interpreter.
\item Wikipedia's list of city latitude/longitudes \cite{wiki}.
\item Software originally written by John Montgomery \cite{tsp} in 2007.General exercise in different types of TSP solving algorithms.
\item The results were then visualized using Google maps static mapping API \cite{google}.

\end{enumerate}

@@

%$yDis = ( lat2 - lat1 ) * \text{Nautical Miles Per Latitude}$\newline
%$xDis = ( \cos( lat1 - \frac{\pi}{180} ) + \cos( lat2 - \frac{\pi}{180} ) ) * ( lon2 - lon1 ) * \frac{\text{Nautical Miles Per Longitude}}{2}$\newline
%$tDistance = \sqrt{ yDis^{2} + xDis^{2} } * \text{Miles per Nautical Miles}$\newline

\begin{figure}[!h]
  \begin{center}
    \setlength\fboxsep{1.00pt}
    \setlength\fboxrule{1.00pt}
    \fbox{\includegraphics[scale=0.50]{path.png}}
    \caption{Final Path Through Africa}
  \end{center}
\end{figure}

Final Order:
\begin{multicols}{3}
\tiny\begin{enumerate*}
\item Congo, Republic of the - Brazzaville
\item Congo, Democratic Republic of the - Kinshasa
\item Angola - Luanda
\item Namibia - Windhoek
\item Botswana - Gaborone
\item South Africa - Pretoria
\item Lesotho - Maseru
\item Swaziland - Mbabane
\item Mozambique - Maputo
\item Zimbabwe - Harare
\item Zambia - Lusaka
\item Malawi - Lilongwe
\item Burundi - Bujumbura
\item Rwanda - Kigali
\item Uganda - Kampala
\item Kenya - Nairobi
\item Tanzania - Dar es Salaam
\item Comoros - Moroni
\item Madagascar - Antananarivo
\item Mauritius - Port Louis
\item Seychelles - Victoria
\item Somalia - Mogadishu
\item Ethiopia - Addis Ababa
\item Djibouti - Djibouti
\item Eritrea - Asmara
\item Sudan - Khartoum
\item Egypt - Cairo
\item Libya - Tripoli
\item Tunisia - Tunis
\item Algeria - Algiers
\item Morocco - Rabat
\item Mauritania - Nouakchott
\item Cape Verde - Praia
\item Senegal - Dakar
\item The Gambia - Banjul
\item Guinea-Bissau - Bissau
\item Guinea - Conakry
\item Sierra Leone - Freetown
\item Liberia - Monrovia
\item Cote d'Ivoire - Yamoussoukro
\item Mali - Bamako
\item Burkina Faso - Ouagadougou
\item Niger - Niamey
\item Ghana - Accra
\item Togo - Lome
\item Benin - Porto-Novo
\item Nigeria - Abuja
\item Chad - N'Djamena
\item Central African Republic - Bangui
\item Cameroon - Yaounde
\item Equatorial Guinea - Malabo
\item Gabon - Libreville
\end{enumerate*}
\end{multicols}

Total Round Trip $\approx 22253.2035387$ miles

\section{Runtime}

The two algorithms, simulated annealing and brute force random permutations where both run in numerious instances over the life of the project. 
In combined total they were run upwards of fity times a peice.

The version of the algorithm wich used simulated annealing on average found the optimal solution in 6 seconds. 
While the bruteforce method often took upwards of

@@

I believe the final optimal path's are within a reasonable uncertenty of the actual path.
Given the big-O of the brute force algorithm $O(n!)$ or $O(80658175170943878571660636856403766975289505440883277824000000000000)$ a runtime complexity of this magnitude is obviously out of my leauge for finding exact values.
At a maxiumum I was able to attempt 100,000,000 permutations of the path.

Two area's of possible improvement I can pinpoint are:
\begin{enumerate}
\item The language itself. I chose the python programming language for my TSP implementation.
The language is interpreted and it's runtime speed can at times be 20\% slower than an equivilant C program.
\item The random.shuffle implementation. Given how python's random.shuffle API is implemented it talkes a set
of values and shuffles them based on a random number between 0,1 generated by the operating system. However given the size of the set
, 52 individual points, the likely hood that random.shuffle will generated all permutations without excessive doubles is increadibly unlikely.
This is the cause of a lot of literally useless computation time, however because of the size of the data set, caching data sets is not plausible let alone if even possible.
\end{enumerate}

Re-writing the algorithm in C would be a great performance benifit. Also having a proper resources to cache routes or a effective random route generater I beleive a more accurate optimal route might be found.

\begin{thebibliography}{99}
\bibitem{python} {\href{ http://python.org }{ http://python.org }}
\bibitem{wiki} {\href{ http://en.wikipedia.org/wiki/Latitude_and_longitude_of_cities }{ http://en.wikipedia.org/wiki/Latitude\_and\_longitude\_of\_cities }}
\bibitem{tsp} Montgomery, John {\it Tackling The Travelling Salesman Problem }{\href{ http://www.psychicorigami.com/category/tsp/ }{ http://www.psychicorigami.com/category/tsp/ } }, {\bf 2007}
\bibitem{google} {\href{ http://code.google.com/apis/maps/documentation/staticmaps/  }{ http://code.google.com/apis/maps/documentation/staticmaps/ }}

\end{thebibliography}
\end{document}
